Analysis and Prediction of Shared Bicycle Usage at the University of Toronto St. George Campus

### 2024-11-25 paper/paper.qmd
adding more on paper

```{r}
#| echo: false
#| include: false
#| warning: false
#| message: false
#### Workspace setup ####
set.seed(912)
# List of necessary packages
packages <- c("tidyverse", "dplyr", "arrow", "sf", "ggplot2", "ggspatial", "RColorBrewer")
# Install missing packages
missing_packages <- packages[!(packages %in% installed.packages()[, "Package"])]
install.packages(missing_packages)
# Load packages
lapply(packages, library, character.only = TRUE)

@@

本研究的核心在于，从自2017年月1日到2024年9月30日中的所有共享单车使用数据中抽取多伦多大学校园内27个站台的样本，并按照4小时的区间，计算每个站台在各个区间内的共享单车使用总量。并且通过贝叶斯柏松回归模型根据站台，年份，月份，日期，以及这一天的具体4小时区间，来预测该站台在此区间内的使用数量。

本研究发现，尽管多大校园内所有站点的使用量每年都呈现极速上涨，但都存在巨大的季节性差异。冬天的使用量无论年份，站点统统呈现着接近0的使用量。另外，在每年的1-9月，使用量呈上升趋势，之后的几个月逐渐下降。一天中不同时间段的使用量也存在明显差异其中8am到8pm的使用量平均占了一天中的79.8%，相比之下0am-8am的使用量只有6.8%。另外值得注意的是27个站台中单车的开出量于停放量基本持平，但存在4个站台他们的进出量纯在较大差异。

这些见解相比于笼统的大多伦多地区共享单车使用数量的统计，更能针对多伦多大学校区内的交通建设以及政策调整，以及如何调度，投放或储存共享单车提供更详细的建议。

结构：TBD

## Estimand
本研究旨在估计特定4小时区间内多伦多大学校园内某个共享单车站台的使用量。通过将时间转换为年，月，日，这一天特定的4小时区间，来考虑到某个站台使用量的总体趋势，季节以及hourly variation。核心目标是探索多伦多大学校园内27个共享单车站台使用量随时间的变化，从而给政策制定者建议以调度自行车，来提高校园内的通勤效率。

@@

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Overview text

@@

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

>>>>

```{r}
#| echo: false
#| include: false
#| warning: false
#| message: false
#### Workspace setup ####
set.seed(912)
# List of necessary packages
packages <- c("tidyverse", "dplyr", "arrow", "sf", "ggplot2", "ggspatial", "RColorBrewer", "knitr")
# Install missing packages
missing_packages <- packages[!(packages %in% installed.packages()[, "Package"])]
install.packages(missing_packages)
# Load packages
lapply(packages, library, character.only = TRUE)

@@

The core of this study is to extract a sample of 27 bike-sharing stations within the University of Toronto campus from all bike-sharing usage data between January 1, 2017, and September 30, 2024. The total usage of bike-sharing at each station is calculated for each 4-hour interval. A Bayesian Poisson regression model is then employed to predict the usage based on the station, year, month, date, and the specific 4-hour interval of the day.

This study found that although the usage at all stations within the University of Toronto campus has increased rapidly each year, there are significant seasonal differences. During winter, usage at all stations, regardless of the year, remains close to zero. Additionally, from January to September each year, the usage shows an upward trend, which then gradually declines over the following months. There are also significant differences in usage across different times of the day. The period from 8 am to 8 pm accounts for an average of 79.8% of daily usage, while the period from midnight to 8 am accounts for only 6.8%. Notably, the number of bikes checked out and parked at the 27 stations is generally balanced; however, there are four stations where there is a substantial discrepancy between the number of check-outs and returns.

Compared to the general statistics on bike-sharing usage in the Greater Toronto Area, these findings offer more detailed recommendations specifically for transportation planning and policy adjustments within the University of Toronto campus. They also provide guidance on how to better allocate, deploy, or store shared bikes.

Structure: TBD

## Estimand
This study aims to estimate the usage of a specific bike-sharing station within the University of Toronto campus during a particular 4-hour interval. By converting time into year, month, day, and the specific 4-hour interval of the day, it accounts for the overall trend, seasonal effects, and hourly variation in station usage. The core objective is to explore the temporal changes in the usage of 27 bike-sharing stations within the campus, thereby providing policymakers with recommendations for bike allocation to improve commuting efficiency on campus.

@@

## Overview
The dataset used in this study comes from opendatatoronto [opendatatoronto], uploaded by Toronto Parking Authority and collected by Bike Share Toronto. It records every bike-sharing usage in the Toronto area from 2015 to September 30, 2024, with a total of 28,017,329 records. The variables included in the data differ across years, but they all contain the following variables: Trip ID, Trip Duration, Trip Start Station ID, Trip Start Time, Trip Start Station Location, Trip End Station ID, Trip End Time, Trip End Station Location, Bike ID, and User Type.

This study follows the workflow of Telling Stories with Data [Telling Stories with Data], using its initial folder structure and part of its code. Data downloading, cleaning, modeling, and visualization were carried out using R [R]. The following R libraries were also used alongside R:

TBD

@@

To predict the relationship between bike-sharing usage and time within the University of Toronto's St. George campus, this study requires a time series to describe the changes in usage at different stations over time. Therefore, the dataset published by opendatatoronto [opendatatoronto], which records every instance of bike-sharing usage in the Toronto area since 2015, is ideal for this study. Additionally, this data is ideal because of its reliability—due to the commercial nature of bike-sharing, the specific time and location of each bike's use and return are accurately recorded. However, the raw data cannot be used directly in this project; sophisticated data cleaning is required, with specific steps and reasons as follows:

First, data from 2015-2016 was excluded because the collection and recording methods for those years differ significantly from later years and do not include the time and station variables required for this study. Next, data for stations located within the University of Toronto's St. George campus was extracted from all remaining samples, and the usage of each station was calculated for every four-hour interval. Finally, the data format was standardized for subsequent analysis. The cleaned data is in the format of @tbl-cleaned_data.

```{r}
#| warning: false
#| message: false
#| echo: false
#| label: tbl-cleaned_data
#| tbl-cap: "Samples of the dataset used for analysis"
start |>
  tail() |>
  kable(booktabs = TRUE, align = "c")
```

## Variables

This study focuses on the following variables:

- `count`: The dependent variable of the study, a non-negative integer. It describes the total usage of bike-sharing at a particular station during a specific 4-hour interval.
- `time`: An independent variable representing the time interval. For example, "2024-09-29 12:00:00" represents the time interval from 12 pm to 4 pm on September 29, 2024, in a 24-hour format. The earliest `time` is "2017-01-01 00:00:00," and the latest is "2024-09-30 24:00:00."
- `station_name`: An independent variable representing the unique name of one of the 27 stations within the University of Toronto's St. George campus.

```{r}
#| label: fig-daily_usage
#| fig-cap: Daily usage from January 2017 to September 2024
#| echo: false
#| message: false
start_asdate <- start 
start_asdate$time <- as.Date(start_asdate$time)
daily_counts <- start_asdate %>%
  group_by(time) %>%
  summarise(total_count = sum(count))
ggplot(data = daily_counts, aes(x = time, y = total_count)) +
  geom_point() +
  labs(x = "Time", y = "Count") +
  theme_minimal()+  
  geom_smooth(method = "lm")  
```

<!-- # Model -->

<!-- The goal of our modelling strategy is twofold. Firstly,... -->

<!-- Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details]. -->

<!-- ## Model set-up -->

<!-- Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.   -->

<!-- \begin{align}  -->
<!-- y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\ -->
<!-- \mu_i &= \alpha + \beta_i + \gamma_i\\ -->
<!-- \alpha &\sim \mbox{Normal}(0, 2.5) \\ -->
<!-- \beta &\sim \mbox{Normal}(0, 2.5) \\ -->
<!-- \gamma &\sim \mbox{Normal}(0, 2.5) \\ -->
<!-- \sigma &\sim \mbox{Exponential}(1) -->
<!-- \end{align} -->

<!-- We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`. -->

<!-- ### Model justification -->

<!-- We expect a positive relationship between the size of the wings and time spent aloft. In particular... -->

<!-- We can use maths by including latex between dollar signs, for instance $\theta$. -->

<!-- # Results -->

<!-- Our results are summarized in @tbl-modelresults. -->

<!-- ```{r} -->
<!-- #| echo: false -->
<!-- #| eval: true -->
<!-- #| warning: false -->
<!-- #| message: false -->

<!-- library(rstanarm) -->

<!-- first_model <- -->
<!--   readRDS(file = here::here("models/first_model.rds")) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- #| echo: false -->
<!-- #| eval: true -->
<!-- #| label: tbl-modelresults -->
<!-- #| tbl-cap: "Explanatory models of flight time based on wing width and wing length" -->
<!-- #| warning: false -->

<!-- modelsummary::modelsummary( -->
<!--   list( -->
<!--     "First model" = first_model -->
<!--   ), -->
<!--   statistic = "mad", -->
<!--   fmt = 2 -->
<!-- ) -->
<!-- ``` -->

<!-- # Discussion -->

<!-- ## First discussion point {#sec-first-point} -->

<!-- If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.  -->

<!-- ## Second discussion point -->

<!-- Please don't use these as sub-heading labels - change them to be what your point actually is. -->

<!-- ## Third discussion point -->

<!-- ## Weaknesses and next steps -->

<!-- Weaknesses and next steps should also be included. -->

<!-- \newpage -->

<!-- \appendix -->

<!-- # Appendix {-} -->

<!-- # Additional data details -->

<!-- # Model details {#sec-model-details} -->

<!-- ## Posterior predictive check -->

<!-- In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows... -->

<!-- In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...  -->

<!-- ```{r} -->
<!-- #| eval: true -->
<!-- #| echo: false -->
<!-- #| message: false -->
<!-- #| warning: false -->
<!-- #| label: fig-ppcheckandposteriorvsprior -->
<!-- #| layout-ncol: 2 -->
<!-- #| fig-cap: "Examining how the model fits, and is affected by, the data" -->
<!-- #| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"] -->

<!-- pp_check(first_model) + -->
<!--   theme_classic() + -->
<!--   theme(legend.position = "bottom") -->

<!-- posterior_vs_prior(first_model) + -->
<!--   theme_minimal() + -->
<!--   scale_color_brewer(palette = "Set1") + -->
<!--   theme(legend.position = "bottom") + -->
<!--   coord_flip() -->
<!-- ``` -->

<!-- ## Diagnostics -->

<!-- @fig-stanareyouokay-1 is a trace plot. It shows... This suggests... -->

<!-- @fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests... -->

<!-- ```{r} -->
<!-- #| echo: false -->
<!-- #| eval: true -->
<!-- #| message: false -->
<!-- #| warning: false -->
<!-- #| label: fig-stanareyouokay -->
<!-- #| fig-cap: "Checking the convergence of the MCMC algorithm" -->
<!-- #| fig-subcap: ["Trace plot", "Rhat plot"] -->
<!-- #| layout-ncol: 2 -->

<!-- plot(first_model, "trace") -->

<!-- plot(first_model, "rhat") -->
<!-- ``` -->

<!-- \newpage -->

<!-- # References -->
